---
title: "EDA-LA1"
output:
  html_document:
    df_print: paged
  pdf_document: default
  word_document: default
date: '2022-06-24'
---

reading the data set csv file
```{r}
mydata<- read.csv("//home//vegita//Downloads//data.csv")
```
1
```{r}
mydata
```
2
to view top of the data
```{r}
head(mydata)
```
3
to view end of the data
```{r}
tail(mydata)
```
4
to view first 10 records
```{r}
head(mydata,10)
```
5
to view last 10
```{r}
tail(mydata,10)
```
6
changing data into string.
```{r}
str(mydata)
```
7
names of the rows
```{r}
names(mydata)
```
8
```{r}
ls()
```

//data frame operations
9
```{r}
x<-data.frame(mydata$Series.Name,mydata$Series.Code,mydata$Country.Name)
```

10
```{r}
x
```
11
column names
```{r}
colnames(mydata)
```

12
rows names
```{r}
rownames(mydata)
```
13
summary of data
```{r}
summary(mydata)
```
14
```{r}
class(mydata)
```
15
```{r}
colSums(is.na.data.frame(mydata))
```

//chunking out portions of data frame 
16
```{r}
first<-(mydata[1:265,])
```
17
```{r}
second<-(mydata[266:529,])
```
18
```{r}
third<-(mydata[530:793,])
```
19

```{r}
fourth<-(mydata[794:1057,])
```

//converting values into numeric values for mathematical operations
```{r}
fourth_num <- as.numeric(mydata[874:886,]$X2016..YR2016.)
```

```{r}
fourth_num1<-as.numeric(fourth$X2016..YR2016.,na.rm=TRUE)
```

```{r}
first_num1=as.numeric(first$X2012..YR2012.,na.rm=TRUE)
```

```{r}
second_num<-as.numeric(second$X2016..YR2016.,na.rm=TRUE)
```

```{r}
first_num<-as.numeric(first$X2016..YR2016.,na.rm=TRUE)
```
```{r}
third_num<-as.numeric(third$X2016..YR2016.,na.rm=TRUE)
```
//finding max

20
```{r}
max(second$X2005..YR2005.)
```

//findin minimum in a column
21
```{r}
min(fourth_num)
```

//Mathematical operations

22
```{r}
median(first$X2005..YR2005.)
```
23
```{r}
mean(fourth_num)
```

24
```{r}
mode(fourth_num)
```




25
```{r}
sum(fourth_num)
```

26
```{r}
sort(fourth_num1)
```
27
```{r}
sort(mydata[1,])
```
28
```{r}

```


//Ranking the collumn values
29
```{r}
rank(fourth$X2010..YR2010.)
```
30
```{r}
dimnames(first)
```

//creating a vectos
31
```{r}
pre=c(rep('open',6),rep('closed',4))
```

```{r}
pre
```

31
```{r}
length(fourth$X2005..YR2005.)
```
// creating a new data frame by selecting columnby choice
32
```{r}
newdata <- data.frame(x1=mydata[794:1057,]$X2010..YR2010., x2=mydata[794:1057,]$X2012..YR2012.)
```

```{r}
newdata
```
33



converting into matrix
34
```{r}
mydata.mat=as.matrix(mydata)
```

```{r}
mydata.mat
```
35
```{r}
frame.list=as.list(first)
```

```{r}
frame.list
```

36
```{r}
vec_new <- mydata[874:886,]$X2010..YR2010.                         
vec_new  
```

caLculating standard deviation
```{r}
sd(vec_new)
```
37
```{r}
vec_new1<-fourth$X2010..YR2010.
```
```{r}
sd(vec_new1,na.rm=TRUE)
```

38
```{r}
length(first$X2010..YR2010.)
```
```{r}
length(na.omit(first$X2010..YR2010.))
```
39
```{r}
log(sd(vec_new1,na.rm=TRUE))
```

40
```{r}
t(first)
```

convertinhg into array
42
```{r}
list1=as.list(fourth$X2010..YR2010.)
arr=array(unlist(list1),dim=c(17,17,1))
print(arr)
```
//table operations

43
```{r}
margin.table(fourth_num)
```

44
```{r}

prop.table(fourth_num)
```

45
```{r}
stem(fourth_num)
```

46
```{r}
data1.norm=rnorm(10,mean(fourth_num),sd(fourth_num))
data1.norm
```

47
```{r}
sample(mydata,size=20,replace=TRUE)
```
48
```{r}
shapiro.test(fourth_num1)
```

49
```{r}
ks.test(first_num,'pnorm',mean=100,sd=27)
```

50
```{r}

wilcox.test(first_num,first_num1);
```

51
```{r}
cor(third_num,fourth_num1)
```
52
```{r}
cov(third_num,fourth_num1)
```
53
```{r}
cor.test(third_num,fourth_num1)
```

56
plotting
```{r}
boxplot(fourth_num1)
```
57
```{r}
boxplot(first_num,third_num,fourth_num1)
```
58
```{r}
plot(fourth_num1)
```
59
```{r}
plot(mydata$X2005..YR2005.,mydata$X2012..YR2012.)
```
60
```{r}
plot(fourth$X2016..YR2016.,type='l')
```

61
```{r}
plot(fourth$X2016..YR2016.,type='p')
```
62
```{r}
plot(fourth$X2016..YR2016.,type='b')
```
63
```{r}
plot(fourth$X2016..YR2016.,type='o')
```

64
```{r}
plot(fourth$X2016..YR2016.,type='c')

```
65
```{r}
plot(fourth$X2016..YR2016.,type='n')
```
66
```{r}
pie(fourth_num,labels = fourth_num)
```
67
```{r}
pie(fourth_num[6])
```
68
```{r}
dotchart(fourth_num1)
```
69
```{r}
dotchart(fourth_num1,color='gray30',gcolor='black',lcolor='gray30',bg='gray90')
```
70
```{r}
barplot(first_num)
```
71
```{r}
barplot(third_num)
```
72
```{r}

second_num
```

73
```{r}
barplot(second_num)
```
74
```{r}
barplot(fourth_num1)

```
75
```{r}
barplot(fourth_num1,xlab='Countries',ylab='Performance Index')
```
76
```{r}
barplot(fourth_num1,xlab='Countries',ylab='Performance Index',ylim=c(0,10),col='lightblue')
```
77
```{r}
barplot(as.matrix(mydata))
```
78
```{r}
plot(first$X2014..YR2014.)
```
79
```{r}
plot(third$X2014..YR2014.)
```
80
```{r}
levels(mydata)
```

82
```{r}
mystack=stack(mydata)
mystack

```
83
```{r}
data2=unstack(mystack)
data2
```
84
```{r}
na.omit(stack(mydata))
```

85
```{r}
x=third$X2016..YR2016.
y=fourth$X2016..YR2016.
x
```

86
binding two vectors
```{r}
cbind(x,y)
```

87
//binding two vectors with rows
```{r}
rbind(x,y)
```



90
```{r}
matrix(x,ncol=4)
```

91
```{r}
first[5]
```

92
sorting 6th column values
```{r}
sort(first[6,])
```
93
//sorting using $ with collumn name
```{r}
sort(first$X2005..YR2005.)
```
94 // making COPY OF DATAFRAME
```{r}
md = mydata
```

95 // searching all packages
```{r}
search()
```
96 // converting dataframe into list
```{r}
md.list = as.list(md)
md.list
```
97.
extracting a individual part of list
```{r}
m1 = md.list$Series.Name
m1
```

99.
finding mean of a vector
```{r}
mean(md.list$Country.Code)
```

100. finding the log of vector
```{r}
cumsum(md.list$X2005..YR2005.)
```
101. creating table
```{r}
table(first_num)
```
102.
creating a data frame from vectors
```{r}
v1  = md.list$Country.Name
v2 = md.list$Country.Code
newdf = data.frame(v1,v2)
newdf
```

